---
title: "2D Feature Tracking Results Analysis"
author: "Ada Lazuli"
date: "2024-01-01"
output:
  pdf_document: default
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(DT)
library(ggplot2)
library(dplyr)
set.seed(101011)
```

## Code Updates and Implementation

### Data Loading

Implementing an input buffer for holding loaded images was the first action item completed for the project. The buffer was implemented as a first in first out structure that holds a limit quantity of images, with that quantity being constrained by a variable set within the application.

When new images are load that would exceed the buffer's specified size, the oldest images are removed from the buffer. The benefit of this being that system resources for data storage are minimized in memory, leaving more resources for processing.

### Detector Implementation

The second phase of the project was it implement a plethora of detectors for keypoint generation on the images loaded in the buffer.

The first detector implemented was __HARRIS__. HARRIS took the most work to implement and utilizes triple layer iteration, which could yield performance issues. Additionally, the HARRIS implementation relied several functions found within openCV 4.1

The remaining detectors, listed below, were implemented strictly using openCV:

1. FAST
2. BRISK
3. ORB
4. AKAZE
5. SIFT

### Input Cropping

The third phase of was to crop the keypoints to a bounding box that contains a singular car. This step is important for later evaluation of detector performance and keypoint matching since the ground truth that the vehicle is the same throughout the images is known.

The cropping feature was implemented by iterating through through the generated keypoints and removing any keypoints that were outside of the bounding box.

### Descriptor Implementation

The fourth phase was to implement several descriptors to process the keypoints. Using functions provided by openCV, the following descriptors were added to the project:

1. BRISK
2. BRIEF
3. ORB
4. FREAK
5. AKAZE
6. SIFT

### Keypoint Matching

The fifth phase was to implement methods for matching processed keypoints between consecutive images. To accomplish this, FLANN and K-Nearest Neighbor matching was implemented in the project.

FLANN was implemented using openCV functions, however, there was a slight complication where varying data types needed to be controlled for. Depending on the descriptor used, there were moments when the data type changed from the expected 32 bit floating point value. This was controlled for by utilizing a conditional to test for the incorrect data type then convert to a float 32 using openCV functions.

Additionally, the K-Nearest Neighbor matcher was implemented by using a the `knnMatch` method of the openCV DescriptorMatcher class.

## Filtering

The sixth and final update to the project was the implmentation of a match filter. The match filter was implemented by iterating over the matched keypoints and evalute the match distance. The match distance would be checked against a specified threshold and used for determining whether to keep or drop the match.

***

## Performance Analysis

### Detector Performance

The purpose of this section is to review the performance of the various detectors implemented in the project and draw conclusions around performance. The quantity of __keypoints__ generated and __duration__ of detection will be the only metrics considered in this section.

The first step is to load the data and overview the comma separated values (CSV) contents.

```{r}
# load the aggregated csv
df <- read.csv("detector_tests.csv")
glimpse(df)
```

The CSV has 60 rows, with three columns per row. The first column details the detector used, the second column records the number of key points generated by the detector for an image, and the third column captures the duration of time needed for keypoint generation.

Additionally, the following code block shows that 6 detectors are uniformly represented in the CSV, with each having 10 associated observations.

```{r}
df %>% count(detector)
```

A box and whisker plot was used to review the distribution of keypoint generation by the various detectors for 10 images. The plot shows that most of the detectors were consistent, with _FAST_ tending to generate the most keypoints. The _HARRIS_ detector was implemented with some custom flow control and exhibited the highest variability.

```{r, out.width="90%"}
ggplot(df, aes(x = detector, y = keypoints)) + geom_boxplot() + xlab("Detector") + 
  ylab("Generated Keypoints") + ggtitle("Comparison Of Keypoint Generation By Detector")
```

Similarly, a box and whisker plot was used to review the distribution of the time used by the detectors for keypoint generation. The plot showed that the _AKAZE_ detector varied the most, _SIFT_ generally took the longest, and _FAST_ was generally the fastest.

```{r, out.width="90%"}
ggplot(df, aes(x = detector, y = duration)) + geom_boxplot() + xlab("Detector") + 
  ylab("Time (ms)") + ggtitle("Duration of Generation By Detector")
```

### Detector Descriptor Performance

A second CSV was created during experimentation that contains details about various runs using different combinations of detectors and descriptors. Additionally, the keypoint matching algorithm and the images used were held constant throughout the experiments.

```{r}
df <- read.csv("detector_descriptor_experiments.csv")
glimpse(df)
```

The CSV contains 315 rows with 6 columns. Each row captures the following details for each detector-descriptor pairing:

1. The __detector__ used for keypoint generation
2. The __descriptor__ used for keypoint description
2. The number of __keypoints__ generated by the detector
3. The amount of __time__ used by the detector
4. The amount of __time__ used by the descriptor
5. The number of keypoint __matches__ between two images

An important aspect to note is that the descriptors are not represented in the data equally. The _AKAZE_ descriptor is only compatible with the _AKAZE_ detector, hence it is represented the least in the data. Additionally, use of the _ORB_ descriptor with the _SIFT_ detector resulted in memory overflow errors and could not be recorded, leaving 9 less examples in the data compared to the other descriptors.

```{r}
df %>% count(descriptor)
```

The matches generated by each detector-descriptor pair were analyzed using a box and whisker plot. _HARRIS_ and _FAST_ had the highest variability and _FAST_ had a tendency for the highest number of keypoint matches. Moreover, the plot showed that the detector played the largest role in performance, compared to the descriptor.

```{r, out.width="90%"}
df <- df %>% mutate(pair = paste(detector, '-', descriptor))
ggplot(df, aes(x = pair, y = matches)) + geom_boxplot() + 
  xlab("Detector-Descriptor Pairing") + ylab("Keypoint Matches") +
  ggtitle("Comparison Of Keypoint Generation By Detector") + coord_flip()
```

A scatter plot was used to further explore the relationship between the number of keypoints and the number of matches. The plot showed that the number of matches increased as the the number of keypoints increased. Therefore, detectors that generate more keypoints, such as _FAST_ would naturally have more matches.

```{r, out.width="90%"}
ggplot(df, aes(x = matches, y = keypoints)) + geom_point(alpha = 0.15) + 
  xlab("Matches") + ylab("Keypoints Generated") + 
  ggtitle("Scatterplot of Keypoints and Matches") + geom_smooth(method='lm')
```

Lastly, the duration for each descriptor was reviewed using a box and whisker plot. The plot showed that the _FREAK_ descriptor had the highest variability, _SIFT_ took the longest, and _BREIF_ was the fastest.

```{r, out.width="90%"}
ggplot(df %>% filter(detector == 'fast'), aes(x = descriptor, y = descriptor.duration)) + 
  geom_boxplot() + xlab("Descriptor") + ylab("Duration (ms)") + 
  ggtitle("Comparison of Descriptor Durations With FAST") + coord_flip()
```

## Conclusion

After reviewing all of the data collected, the top three pairings of detector and descriptor would to be:

1.  __FAST__ with __BRIEF__.
2. __FAST__ with __ORB__.
3. __FAST__ with __BRISK__.


Overall, using the __FAST__ with __BRIEF__ is the best pairing of detector and descriptor. This combination results in the highest keypoint generation, matching, and shortest duration. 

