---
title: "2D Feature Tracking Results Analysis"
author: "Ada Lazuli"
date: "2024-01-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(DT)
library(ggplot2)
library(dplyr)
set.seed(101011)
```

## Part 1 - Detector Performance

The purpose of this section is to review the performance of the various detectors implemented in the project and draw conclusions around performance. The quantity of __keypoints__ generated and __duration__ of detection will be the only metrics considered in this section.

The first step is to load the data and overview the comma separated values (CSV) contents.

```{r}
# load the aggregated csv
df <- read.csv("detector_tests.csv")
glimpse(df)
```

The CSV has 60 rows, with three columns per row. The first column details the detector used, the second column records the number of key points generated by the detector for an image, and the third column captures the duration of time needed for keypoint generation.

Additionally, the following code block shows that 6 detectors are uniformly represented in the CSV, with each having 10 associated observations.

```{r}
df %>% count(detector)
```

A box and whisker plot was used to review the distribution of keypoint generation by the various detectors for 10 images. The plot shows that most of the detectors were consistent, with _FAST_ tending to generate the most keypoints. The _HARRIS_ detector was implemented manually and exhibited the highest variability.

```{r}
ggplot(df, aes(x = detector, y = keypoints)) + geom_boxplot() + xlab("Detector") + ylab("Generated Keypoints") + ggtitle("Comparison Of Keypoint Generation By Detector")
```

Similarly, a box and whisker plot was used to review the distribution of the time used by the detectors for keypoint generation. The plot showed that the _AKAZE_ detector varied the most, _SIFT_ generally took the longest, and _FAST_ was generally the fastest.

```{r}
ggplot(df, aes(x = detector, y = duration)) + geom_boxplot() + xlab("Detector") + ylab("Time (ms)") + ggtitle("Duration of Generation By Detector")
```

## Part 2 - Detector Descriptor Performance

A second CSV was created during experimentation that contains details about various runs using different combinations of detectors and descriptors. Additionally, the keypoint matching algorithm and the images used were held constant throughout the experiments.

```{r}
df <- read.csv("detector_descriptor_experiments.csv")
glimpse(df)
```

The CSV contains 315 rows with 6 columns. Each row captures the following details for each detector-descriptor pairing:

1. The __detector__ used for keypoint generation
2. The __descriptor__ used for keypoint description
2. The number of __keypoints__ generated by the detector
3. The amount of __time__ used by the detector
4. The amount of __time__ used by the descriptor
5. The number of keypoint __matches__ between two images

An important aspect to note is that the descriptors are not represented in the data equally. The _AKAZE_ descriptor is only compatible with the _AKAZE_ detector, hence it is represented the least in the data. Additionally, use of the _ORB_ descriptor with the _SIFT_ detector resulted in memory overflow errors and could not be recorded, leaving 9 less examples in the data compared to the other descriptors.

```{r}
df %>% count(descriptor)
```

The matches generated by each detector-descriptor pair were analyzed using a box and whisker plot. _HARRIS_ and _FAST_ had the highest variability and _FAST_ had a tendency for the highest number of keypoint matches. Moreover, the plot showed that the detector played the largest role in performance, compared to the descriptor.

```{r}
df <- df %>% mutate(pair = paste(detector, '-', descriptor))
ggplot(df, aes(x = pair, y = matches)) + geom_boxplot() + xlab("Detector-Descriptor Pairing") + ylab("Keypoint Matches") + ggtitle("Comparison Of Keypoint Generation By Detector") + coord_flip()
```

A scatter plot was used to further explore the relationship between the number of keypoints and the number of matches. The plot showed that the number of matches increased as the the number of keypoints increased. Therefore, detectors that generate more keypoints, such as _FAST_ would naturally have more matches.

```{r}
ggplot(df, aes(x = matches, y = keypoints)) + geom_point(alpha = 0.5) + xlab("Matches") + ylab("Keypoints Generated") + ggtitle("Scatterplot of Keypoints and Matches") + geom_smooth(method='lm')
```

Lastly, the duration for each descriptor was reviewed using a box and whisker plot. The plot showed that the _FREAK_ descriptor had the highest variability, _SIFT_ took the longest, and _BREIF_ was the fastest.

```{r}
ggplot(df %>% filter(detector == 'fast'), aes(x = descriptor, y = descriptor.duration)) + geom_boxplot() + xlab("Descriptor") + ylab("Duration (ms)") + ggtitle("Comparison of Descriptor Durations With FAST") + coord_flip()
```

## Conclusion

After reviewing all of the data collected, the best pairing of detector and descriptor appears to be __FAST__ with __BRIEF__. The combination results in the highest keypoint generation, matching, and shortest duration. 

